On hyperparameter tuning through Lipschitz global optimization

### 2024-08-03 2024/ITMO paper/sn-article.tex
Rev1 submitted

\textcolor[rgb]{0,0,1}{
\begin{remark}
In this work we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}). It is simple from the point of view of implementation and use, but does not reflect all the properties of the Peano-Hilbert curve. More complex approximations fit the curve better. For example, in \cite{Strongin2000} a non-univalent approximation was also proposed. It reflects the following property of the Peano-Hilbert curve $y(x)$: a point in $D$ can have several (but not more than $2^N$) inverse images in $[0,1]$. Using non-univalent approximation allows one to reduce the number of objective function evaluations required to solve the problem with a given accuracy (see \cite{Sergeyev2022, Lera2024}).
\end{remark}}

@@

Let us now consider the mapping
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
which maps any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
\begin{equation}\label{f}
f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
\end{equation}
which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
\textcolor[rgb]{0,0,1}{For the same treatment of boundary points of the interval $[0,S]$ we will also consider the values $f(0)$ and $f(S)$ to be undefined. The function values at boundary points will not be used in the algorithm.}

@@

The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
\textcolor[rgb]{0,0,1}{In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods. In such problems there is a lack of analytical description of the objective function and the function itself can be multi-extremal.}

@@

\begin{enumerate}[1.]
\item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
\textcolor[rgb]{0,0,1}{Recall that the values $z^i$ at integer points $x^i=i, 0\leq i \leq S,$ are supposed to be undefined and are not used later in the algorithm.}
After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information. Thus, after initialization, the set $\Omega_k$ will contain $k=2S+1$ elements.

>>>>

\begin{remark}
In this work, we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}). This approximation is simple from the point of view of implementation and use, but it does not reflect all the properties of the Peano-Hilbert curve. More complex approximations fit the curve better. For example, in \cite{Strongin2000} a non-univalent approximation was also proposed. It reflects the following property of the Peano-Hilbert curve $y(x)$: a point in $D$ can have several (but not more than $2^N$) inverse images in $[0,1]$. Using non-univalent approximation allows one to reduce the number of objective function evaluations required to solve the problem with a given accuracy (see \cite{Sergeyev2022, Lera2024}).
\end{remark}

@@

Let us now consider the mapping
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
which maps any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
\begin{equation}\label{f}
f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
\end{equation}
which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
For treating the boundary points of the interval $[0,S]$ in a uniform manner, we will also consider the values $f(0)$ and $f(S)$ to be undefined. The function values at boundary points will not be used in the algorithm.

@@

The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods. In such problems, there is a lack of analytical description of the objective function and the function itself may be multi-extremal.

@@

\begin{enumerate}[1.]
\item At the initialization stage, the values $(x^i = i, \nu^i=0, z^i)$, $ 0\leq i \leq S$, i.e., the boundary points of the intervals $(x_{i-1},x_i), 1<i\leq S$, are entered into the set $\Omega_k$ containing the search  information.
Recall that the values $z^i$ at integer points $x^i=i, 0\leq i \leq S,$ are assumed to be undefined and are not used later in the algorithm.
After that, trials are carried out at arbitrary internal points (e.g., midpoints) of the intervals $(x_{i-1},x_i), 1<i\leq S$; the trial results are also added to the search information. Thus, after initialization, the set $\Omega_k$ will contain $k=2S+1$ elements.
